Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} Model System

### 2014-04-25 trunk/paper_enahpc14/Conclusion.tex
Manuel's changes

We  have   presented  a  methodology  for   comparing  performance  of
COSMO-ART,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: an Intel Sandybridge (2.6 GHz) system has
lower  time-to-solution but  higher energy-to-solution  than  an Intel
Ivybridge  (2.2  GHz) system,  although  the  two  metrics are  indeed
strongly  correlated.   A   tool  utilising  Paraver/Extrae  with  new
extensions  for power  consumption  was successfully  applied to  this
non-trivial application.  The  resulting profiles indicate that simple
changes, such as making use of an MPI version based on blocking rather
than polling, can reduce power consumption significantly.

>>>>

We  have   presented  a  methodology  for   comparing  performance  of
COSMO-ART,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: an Intel Sandybridge (2.6 GHz) system has
lower  time-to-solution but  higher energy-to-solution  than  an Intel
Ivybridge  (2.2  GHz) system,  although  the  two  metrics are  indeed
strongly  correlated. On the other hand, the usage of energy-friendly MPI waiting techniques analzyed with our power-performance framework over a series of runs of \cosmoart on \tinto and
power traces demonstrate that it is possible to reduce both
power dissipation and energy-to-solution while maintaining (or even decreasing) the time-to-solution. The  resulting profiles indicate that simple
changes, such as making use of the blocking MPI policy rather
than the polling is possible to slightly reduce both power dissipation and energy consumption (by 2\,\%).

### 2014-04-25 trunk/paper_enahpc14/Introduction.tex
Manuel's changes

While anthropogenic greenhouse gas emissions are driving unprecedented
major  climate  changes  since   the  mid-20$^{th}$  Century,  one  factor
overwhelmingly  affects  the  uncertainty  in  defining  human-induced
radiative  forcing:  the  effects   of  aerosols  in  the  atmosphere.
Although they  are not considered  heat-trapping greenhouse  gases and
have  shorter  atmospheric lifetimes,  aerosols  significantly modify  the
global   radiation   budget   \citep{IPCC-2013}.   Enhanced   aerosols
concentrations impact the climate  system by scattering and absorption
of solar radiation, thereby exerting a direct radiative forcing; or by
modifying cloud  properties,  cloud  fraction and  surface
albedo,  causing  a  negative  indirect  radiative  forcing.   Despite
considerable  progress in  global aerosol  modelling \citep{Mann-2013}
and  measurement-based  assessments,  large  uncertainties  remain  in
current  estimates  of  aerosol radiative  forcing  \citep{Myhre-2013,
IPCC-2013,  Lee-2013,   Randles-2013,  Rosenfeld-2013,  Sherwood-2013,
Stier-2013}.

Hence to  improve our understanding of  aerosol-cloud interactions and
reduce  these  uncertainties,  the  research  community  is  making  a
concerted international  effort to represent  the underlying physical,
chemical  and aerosol dynamical  processes through  numerical chemical
transport  models (CTMs)  such  as ART  (Aerosols  and Reactive  Trace
gases),   developed   at  the   Karlsruhe   Institute  of   Technology
(KIT)  \citep{Vogel-2009,  Bangert-2011,  Knote-2013}.  This  regional
scale modelling system is coupled  with the operational  weather forecast
model  \textsc{Cosmo}  \citep{Baldauf-2011},  jointly developed  by  a
consortium of European weather centers and also used in a climate version
by the wider research community.

The extended  atmospheric model \textsc{Cosmo-art}  is com\-put\-ationally
much  more  demanding than  \textsc{Cosmo}  since  a  large number  of
additional  tracers and  processes have  to be  considered.   Thus the
model  is currently  severely-limited  in terms  of applicability  and
expensive in terms of energy consumption.  Although \textsc{Cosmo} has
recently been ported to GPUs \citep{Gysi-2014, Lapillonne-2014} within
the framework of the  High Performance and High Productivity Computing
(HP2C) Initiative  \citep{HP2C} to optimise it  for computational and
energy efficiency,  significant investments in ART  are still required
to take it  to a similar level.  The  efficiency of \textsc{Cosmo-art}
is being  addressed in the  EU Exa2Green project  \citep{EXA2GREEN} to
deliver a  prototype code, which  provides an energy efficiency  of at
least five times of the  baseline value.  Such an implementation would
allow  the  community  to  investigate critical  questions  at  higher
resolution  and   over  longer  periods,   at  reduced  cost   to  the
environment.

This  work is  organized as  follows: in  Sec.~\ref{sec:1} we  give an
overview  of  related  work,   then  in  Sec.~\ref{sec:2}  we  briefly
introduce \textsc{Cosmo-art}  and specify its  technical setup related
to  the investigated  performance and  energy evaluation  methods.  In
Sec.~\ref{sec:3}   we  describe   HPC  platforms,   power  measurement
equipments   and  software  environment   employed  to   conduct  this
benchmarking study.  Sec.~\ref{sec:4}  presents performance and energy
requirements  of the  baseline on  these architectures  and highlights
areas where improvements will be necessary for the subsequent baseline
refactoring.   Finally,  we  conclude in Sec.~\ref{concl} with some 
implications for the Exa2Green project and give an outlook for future
research.

>>>>

While anthropogenic greenhouse gas emissions are driving unprecedented
major  climate  changes  since   the  mid-20$^{th}$  Century,  one  factor
overwhelmingly  affects  the  uncertainty  in  determining  human-induced
radiative  forcing:  the  effects   of  aerosols  in  the  atmosphere.
Although they  are not considered  heat-trapping greenhouse  gases and
have  shorter  atmospheric lifetimes,  aerosols  significantly modify  the
global   radiation   budget   \cite{IPCC-2013}.   Enhanced   aerosols
concentrations impact the climate  system by scattering and absorption
of solar radiation, thereby exerting a direct radiative forcing; or by
modifying cloud  properties,  cloud  fraction and  surface
albedo,  causing  a  negative  indirect  radiative  forcing.   Despite
considerable  progress in  global aerosol  modelling \cite{Mann-2013}
and  measurement-based  assessments,  large  uncertainties  remain  in
current  estimates  of  aerosol radiative  forcing  \cite{Myhre-2013,
IPCC-2013, Lee-2013,  Sherwood-2013,  Stier-2013}.

Hence to  improve our understanding of  aerosol-cloud interactions and
reduce  these  uncertainties,  the  research  community  is  making  a
concerted international  effort to represent  the underlying physical,
chemical  and aerosol dynamical  processes through  numerical chemical
transport  models (CTMs)  such  as ART  (Aerosols  and Reactive  Trace
gases),   developed   at  the   Karlsruhe   Institute  of   Technology
(KIT)  \cite{Vogel-2009,  Bangert-2011,  Knote-2013}.  This  regional
scale modelling system is coupled  with the operational  weather forecast
model  \textsc{Cosmo}  \cite{Baldauf-2011},  jointly developed and used by  a
consortium of European weather centers, as well as utilized in a climate version
by the wider research community.

The extended  atmospheric model \cosmoart  is com\-put\-ationally
much  more  demanding than  \textsc{Cosmo}  since  a  large number  of
additional  tracers and  processes have  to be  considered.   Thus the
model  is currently  severely limited  in terms  of applicability  and
expensive in terms of energy consumption.  Although \textsc{Cosmo} has
recently been ported to GPUs \cite{Gysi-2014, Lapillonne-2014} 
within
the framework of the  High Performance and High Productivity Computing
(HP2C) Initiative  \cite{HP2C} to optimize it  for computational and
energy efficiency,  significant investments in ART  are still required
to take it  to a similar level.  The  efficiency of \cosmoart
is being  addressed in the  EU Exa2Green project  \cite{EXA2GREEN} to
deliver a  prototype code, which  provides an energy efficiency  of at
least five times the  baseline value.  Such an implementation would
allow  the  community  to  investigate critical  questions  at  higher
resolution  and   over  longer  periods,   at  reduced  cost   to  the
environment.

This  work is  organized as  follows: in  Section~\ref{sec:1} we  give an
overview  of  related  work,   then  in  Section~\ref{sec:2}  we  briefly
introduce \cosmoart  and specify its  technical setup related
to  the investigated  performance and  energy evaluation  methods.  In
Section~\ref{sec:3}   we  describe   HPC  platforms,   power  measurement
equipments   and  software  environment   employed  to   conduct  this
benchmarking study.  Section~\ref{sec:4}  presents performance and energy
requirements  of the  baseline on  these architectures  and highlights
areas where improvements will be necessary for the subsequent baseline
refactoring.   Finally,  we  conclude in Section~\ref{concl} with some 
implications for the Exa2Green project and give an outlook for future
research.

### 2014-04-25 trunk/paper_enahpc14/Sec3:Environment-setup.tex
Manuel's changes

\begin{itemize}
\item \monch is a 10-rack NEC-provided clus\-ter at CSCS composed
  of 312 compute  nodes. %and utilised by members  of the Swiss Platform
  %for Advanced Scientific  Computing (PASC) project~\citep{PASC}
  Each node is comprised of two Intel Ivy Bridge EP  E5-2660 ten-core processors
  operating at  2.2 GHz, equipped with 32GB of DDR3 1600\,MHz  RAM and connected  via InfiniBand Mellanox  SX6036 and FDR switches. For our  experiments, 
  a full rack constituted of 52 nodes was used. % \monch is  slated to stay
  %in  service  without  hardware  upgrade  for  the  duration  of  the
  %Exa2Green  project,  allowing   an  identical  configuration  to  be
  %recreated for future assessments of the baseline.\\

\item \pilat is a dual-socket eight-core Intel Sandy
  Bridge EP based cluster at  CSCS composed of 42 compute nodes. %and used as Piz Daint pre-post  processing cluster 
  %The 2 login and 42 compute nodes %are consolidated into 11 twin-pair Intel E5-Series DALCO
  %r2264i4t 2U scalable compute modules, each containing 4 compute
  %nodes based 
  %consisting of 
  Each node is comprises two Intel Xeon E5-2670 processors operating  at 2.6\,GHz equipped with 64GB of DDR3 1600\,MHz RAM and connected  via InfiniBand Mellanox SX6036 and FDR switches. 
  For our  experiments, a full rack constituted of 42 compute nodes
  was used. \pilat  is based  on  Intel's second-generation
  processors, conventional in  HPC systems and known to  be more power
  consuming than its Ivy Bridge successor.\\

\item \tinto is  a heterogeneous cluster composed of
  28 compute nodes at UJI. For our  experiments only a subset of 16 homogeneous nodes was considered. Each node is composed of two Intel Xeon E5645 hexa-core processors 
  running at 2.4\,GHz, equipped with 24\,GB  of DDR3 1333\,MHz and connected via Infiniband QDR with a Mellanox MTS\-3600 switch.

@@

Supercomputer  clusters considered  for our  experiments at CSCS of ETH Zurich are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high   quality   electricity  meters   released   by  Riedo   Networks
\citep{Riedonetworks},   that  enable   to  monitor   and   log  power
dissipation  of the IT  infrastructure as  well as  constantly analyse
line  voltage,  current, power-factor,  frequency  with 1\,\%  accuracy.
Using  reliable narrowband  powerline communication  (PLC) technology,
all  metering and  power  quality  data from  each  IPS are  centrally
collected  by the E3METER  Data Concentrator,  via the  existing power
cables thus  avoiding the need for  extra cabling.  This  data is made
available via  SNMP, HTTP and TELNET  through the built-in  Fast Ethernet
port.   Time  synchronisation  is  guaranteed by  using  NTP  servers.
Measured  data is  accessed  through the  open  source Cacti  software
including the E3METER Cacti Plugin  to scan the entire PLC network and
monitor in real-time the power usage of individual rack, recorded in 5
minutes interval periods.

@@

For experiments at UJI on \tinto, we employed a version
of the integrated framework presented in~\cite{energy13} that employs the \pmlib library in
combination with Extrae and Paraver, which  are profiling/tracing and
visualisation tools, respectively.

To employ our framework, \cosmoart is first compiled and linked with the Extrae instrumentation library, which automatically  instruments the Fortran~90 code of the model, e.g., to record MPI calls.  Next, \cosmoart  is  run on  the  nodes,  thus dissipating  certain amount  of power. The  nodes are connected to power distribution units (PDUs)
APC 8653 that account for the dissipated power with a sampling rate of 1\,Hz. The power data captured by these PDUs is collected by a \pmlib tracing server running on a separate machine. The client, running on the target nodes, sends start/stop  primitive in order to collect
the power data during the \cosmoart run. Once  the run  is finished,  a power-performance trace file containing the power data is received from the tracing server and the instrumentation post-process is generated. A merged trace file is finally loaded in Paraver to visualise and correlate the performance trace and with the power profile generated by \cosmoart. This procedure allows to detect power bottlenecks, basically constituted by inefficient waiting methods.

@@

The \cosmoart  baseline is a  pure MPI-based  Fortran~90 code
currently  running  on   distributed  multi-core  systems  only.   The
software stack on the platforms was controlled using the modules
framework which gives an easy  and flexible mechanism to access to all
of the  provided compilers, tools  and applications.  For  our initial
benchmarking,   we  opted   for   the  GNU   compiler  GCC 4.8.1   on
\monch and GCC 4.8.2 on  \pilat and \tinto) using the -O3 compiler
flag  in favor  of the  Intel compilers 14.0.1,  which delivered inferior
performance.  In  addition, we installed the following implementations of MPI: MVAPICH2 1.9 on \monch and \pilat and  OpenMPI 1.6.5 on \tinto. On the other hand, we  used HDF5 1.8.12 and NetCDF  (4.3.1) libraries for  the management  of extremely  large and
complex data collections.  All computes nodes have an operating system
based on GNU/Linux featuring ``2.6.32-358.11.1.el6'' kernel in
\monch, ``3.0.101-0.15-default'' in \pilat and ``2.6.18-238.19.1.el5'' in \tinto.

A snapshot of the code,  which includes, at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
\cosmoart, was produced and run on 1040 cores using 20~MPI
tasks  per node on  \monch, on 1344 cores  using 16~MPI
tasks per node on  \pilat and on 192 cores  using 12~MPI
tasks per node on \tinto.  The calculated region was mapped
to the participating processors using a 2D-partitioning strategy.  The
distribution along the $x$ and $y$ coordinates was defined accordingly, keeping in mind
% $by setting:
%$nprocx=40$  and $nprocy=26$  for \monch  and  $nprocx=28$ and
%$nprocy=24$  as  
to keep $x$  bigger  than  $nprocy$.
Besides,  as this  version does not  make use  of the  traditional GRIB
library, we  specified $nprocio=0$  for GRIB I/O.   Hyperthreading was
not considered in this study  as it previously revealed that it always
led to higher energy-to-solution.

>>>>

\begin{itemize}
\item \monch is a 10-rack NEC-provided clus\-ter at CSCS composed
  of 312 compute  nodes. %and utilized by members  of the Swiss Platform
  %for Advanced Scientific  Computing (PASC) project~\cite{PASC}
  Each node is comprised of two Intel Ivy Bridge EP  E5-2660 ten-core processors
  operating at  2.2\,GHz, equipped with 32\,GB of DDR3 1600\,MHz  RAM and connected  via InfiniBand Mellanox  SX6036 and FDR switches. For our  experiments, 
  a full rack constituted of 52 nodes was used. % \monch is  slated to stay
  %in  service  without  hardware  upgrade  for  the  duration  of  the
  %Exa2Green  project,  allowing   an  identical  configuration  to  be
  %recreated for future assessments of the baseline.\\

\item \pilat is a dual-socket eight-core Intel Sandy
  Bridge EP based cluster at  CSCS composed of 42 compute nodes. %and used as Piz Daint pre-post  processing cluster 
  %The 2 login and 42 compute nodes %are consolidated into 11 twin-pair Intel E5-Series DALCO
  %r2264i4t 2U scalable compute modules, each containing 4 compute
  %nodes based 
  %consisting of 
  Each node is comprises two Intel Xeon E5-2670 processors operating  at 2.6\,GHz equipped with 64\,GB of DDR3 1600\,MHz RAM and connected  via InfiniBand Mellanox SX6036 and FDR switches. 
  For our  experiments, a full rack constituted of 42 compute nodes
  was used. \pilat  is based  on  Intel's second-generation
  processors, conventional in  HPC systems and known to  be more power
  consuming than its Ivy Bridge successor.

\item \tinto is  a heterogeneous cluster composed of
  28 compute nodes at UJI. For our  experiments only a subset of 16 homogeneous nodes was considered. Each node is composed of two Intel Xeon E5645 hexa-core processors 
  running at 2.4\,GHz, equipped with 24\,GB  of DDR3 1333\,MHz and connected via InfiniBand QDR with a Mellanox MTS\-3600 switch.

@@

Supercomputer  clusters considered  for our  experiments at CSCS of ETH Zurich are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high   quality   electricity  meters   released   by  Riedo   Networks, 
  that  enable   to  monitor   and   log  power
dissipation  of the IT  infrastructure as  well as  constantly analyze
line  voltage,  current, power-factor,  frequency  with 1\,\%  accuracy.
Using  reliable narrowband  power-line communication  (PLC) technology,
all  metering and  power  quality  data from  each  IPS are  centrally
collected  by the E3METER  Data Concentrator,  via the  existing power
cables thus  avoiding the need for  extra cabling.  This  data is made
available via  SNMP, HTTP and TELNET  through the built-in  Fast Ethernet
port.   Time  synchronization  is  guaranteed by  using  NTP  servers.
Measured  data is  accessed  through the  open  source Cacti  software
including the E3METER Cacti Plugin  to scan the entire PLC network and
monitor in real-time the power usage of individual rack, recorded in 5
minutes interval periods.

@@

For experiments at UJI on \tinto, we employed a version
of the integrated framework presented in~\cite{energy13} that employs the \pmlib library in
combination with Extrae and Paraver, which  are profiling/tracing and
visualization tools, respectively.

To employ our framework, \cosmoart is first compiled and linked with the Extrae instrumentation library, which automatically  instruments the Fortran~90 code of the model, e.g., to record MPI calls.  Next, \cosmoart  is  run on  the  nodes,  thus dissipating  certain amount  of power. The  nodes are connected to APC 8653 Power Distribution Units (PDUs)
that account for the dissipated power with a sampling rate of 1\,Hz. The power data captured by these PDUs is collected by a \pmlib tracing server running on a separate machine. The client, running on the target nodes, sends start/stop  primitives in order to collect
the power data during the \cosmoart run. Once  the run  is finished,  a power-performance trace file containing the power data is received from the tracing server and the instrumentation post-process is generated. A merged trace file is finally loaded in Paraver to visualize and correlate the performance trace and with the power profile generated by \cosmoart. This procedure allows us to easily detect power bottlenecks, basically constituted by inefficient waiting methods.

@@

The \cosmoart  baseline is a  pure MPI-based  Fortran~90 code
currently  running  on   distributed  multi-core  systems  only.   The
software stack on the platforms was controlled using the modules
framework which gives an easy  and flexible mechanism to access to all
of the  provided compilers, tools  and applications.  For  our initial
benchmarking,   we  opted   for   the  GNU   compiler  GCC 4.8.1   on
\monch and GCC 4.8.2 on  \pilat and \tinto) using the -O3 compiler
flag  in favor  of the  Intel compilers 14.0.1,  which delivered inferior
performance.  In  addition, we installed the following implementations of MPI: MVAPICH2 1.9 on \monch and \pilat, and  OpenMPI 1.6.5 on \tinto. On the other hand, we  used HDF5 1.8.12 and NetCDF  (4.3.1) libraries for  the management  of extremely  large and
complex data collections.  All computes nodes have an operating system
based on GNU/Linux featuring ``2.6.32-358.11.1.el6'' kernel in
\monch, ``3.0.101-0.15-default'' in \pilat and ``2.6.18-238.19.1.el5'' in \tinto.

A snapshot of the code,  which includes, at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
\cosmoart, was produced and run on 1040 cores using 20~MPI
tasks  per node on  \monch, on 1344 cores  using 16~MPI
tasks per node on  \pilat and on 192 cores  using 12~MPI
tasks per node on \tinto.  The calculated region was mapped
to the participating processors using a 2D-partitioning strategy.  The
distribution along the $x$ and $y$ coordinates was defined accordingly depeding on the selected platform.
% $by setting:
%$nprocx=40$  and $nprocy=26$  for \monch  and  $nprocx=28$ and
%$nprocy=24$  as  
%to keep $x$  bigger  than  $nprocy$.
Besides,  as this  version does not  make use  of the  traditional GRIB
library, we  specified $nprocio=0$  for GRIB I/O.   Hyperthreading was
not considered in this study  as it previously revealed that it always
led to higher energy-to-solution.

### 2014-04-25 trunk/paper_enahpc14/paper.tex
Manuel's changes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
\documentclass[twocolumn]{svjour3}
%
\smartqed
%
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage[round, sort&compress, numbers]{natbib}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{units}
\usepackage{color}
\usepackage{colortbl}
\usepackage{xspace}
\usepackage{epsfig}
%
\DeclareRobustCommand\IPCClongname{}
\setcounter{secnumdepth}{5}
%
% please place your own definitions here and don't use \def but \newcommand{}{}
\newcommand{\pmlib}{\texttt{pmlib}\xspace}
\newcommand{\monch}{\textsc{Monch}\xspace}
\newcommand{\pilat}{\textsc{Pilatus}\xspace}
\newcommand{\tinto}{\textsc{Tintorrum}\xspace}
\newcommand{\cosmoart}{\textsc{Cosmo-Art}\xspace}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Evaluating the Performance and Energy Efficiency of the
  \textsc{COSMO-ART} model system}
%\thanks{}
%\subtitle{}

@@

\begin{abstract}
  In this  paper we investigate  the energy footprint  and performance
  profiling  of  \textsc{COSMO-ART} on  various  HPC platforms.   This
  model is an  extension of the operational weather  forecast model of
  the German  weather service (DWD),  developed for the  evaluation of
  the interactions  of reactive gases  and aerosol particles  with the
  state of  atmosphere at  the regional scale.   Different measurement
  devices and  energy-aware techniques are described  to evaluate both
  time and  energy to  solution of the  considered application  and to
  gain detailed insights into power and performance requirements.  Our
  motivation  is to  improve  corresponding code  sections to  sustain
  performance  while minimising energy-to-solution.   This preliminary
  work  sets the  basis for  subsequent studies  to  tackle challenges
  related  to  energy  efficient  high performance  computing  in  the
  framework of the Exa2Green project \citep{EXA2GREEN}.

\keywords{High performance computing  \and Energy-aware computing \and
  Green computing  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosol  modelling  \and  Profiling  methods  \and
  Benchmark  analysis  \and   \textsc{COSMO-ART}  coupled  model  \and
  Time-to-solution \and Energy-to-solution \and Xeon processor}
% \PACS{PACS code1 \and PACS code2 \and more} 
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

\section{Related work}
\label{sec:1}
\input{Sec1:Related.tex}

\section{The \textsc{COSMO-ART} model system}
\label{sec:2}
\input{Sec2:COSMO-ART.tex}

\section{Environment setup}
\label{sec:3}
\input{Sec3:Environment-setup.tex}

\section{Experimental results}
\label{sec:4}
\input{Sec4:Experimental-results.tex}

\section{Conclusion}
\label{concl}
\input{Conclusion.tex}

@@

\begin{acknowledgements}
The research  leading to these  results is supported by  the Exa2Green
project  co-financed by  the European  Commission under  7th Framework
Programme Future and Emerging Technologies (FET) Proactive Initiative:
Minimising Energy  Consumption of Computing to the  Limit (MINECC). We
also gratefully acknowledge the High Performance and High Productivity
Computing Initiative  \citep{HP2C} for results that  will be leveraged
in subsequent code refactoring.
\end{acknowledgements}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
\documentclass[twocolumn]{svjour3}
%
\smartqed
%
\usepackage[numbers]{natbib}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{colortbl}
\usepackage{url}
\usepackage{xspace}

%
\DeclareRobustCommand\IPCClongname{}
\setcounter{secnumdepth}{5}
%
% please place your own definitions here and don't use \def but \newcommand{}{}
\newcommand{\pmlib}{\texttt{pmlib}\xspace}
\newcommand{\monch}{\textsc{Monch}\xspace}
\newcommand{\pilat}{\textsc{Pilatus}\xspace}
\newcommand{\tinto}{\textsc{Tintorrum}\xspace}
\newcommand{\cosmoart}{\textsc{Cosmo-Art}\xspace}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Evaluating the Performance and Energy Efficiency of the
  \textsc{COSMO-ART} Model System}
%\thanks{}
%\subtitle{}

@@

\begin{abstract}
  In this  paper we investigate  the energy footprint  and performance
  profiling  of  \textsc{COSMO-ART} on  various  HPC platforms.   This
  model is an  extension of the operational weather  forecast model of
  the German  weather service (DWD),  developed for the  evaluation of
  the interactions  of reactive gases  and aerosol particles  with the
  state of  atmosphere at  the regional scale.   Different measurement
  devices and  energy-aware techniques are described  to evaluate both
  time and  energy to  solution of the  considered application  and to
  gain detailed insights into power and performance requirements.  Our
  motivation  is to  improve  corresponding code  sections to  sustain
  performance  while minimizing energy-to-solution.   This preliminary
  work  sets the  basis for  subsequent studies  to  tackle challenges
  related  to  energy  efficient  high performance  computing  in  the
  framework of the Exa2Green project \cite{EXA2GREEN}.

\keywords{High performance computing  \and Energy-aware computing
  %Green computing  
  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosol  modelling \and
  Benchmark  analysis  \and 
  \cosmoart  coupled  model \and  
  Power-performance profiling and tracing tools  \and
  Energy-saving techniques}
% \PACS{PACS code1 \and PACS code2 \and more} 
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

\section{Related Work}
\label{sec:1}
\input{Sec1:Related.tex}

\section{The \textsc{COSMO-ART} Model System}
\label{sec:2}
\input{Sec2:COSMO-ART.tex}

\section{Environment Setup}
\label{sec:3}
\input{Sec3:Environment-setup.tex}

\section{Experimental Results}
\label{sec:4}
\input{Sec4:Experimental-results.tex}

\section{Concluding Remarks}
\label{concl}
\input{Conclusion.tex}

@@

\begin{acknowledgements}
The research  leading to these  results is supported by  the Exa2Green
project  co-financed by  the European  Commission under  7th Framework
Programme Future and Emerging Technologies (FET) Proactive Initiative:
Minimizing Energy  Consumption of Computing to the  Limit (MINECC). We
also gratefully acknowledge the High Performance and High Productivity
Computing Initiative  \cite{HP2C} for results that  will be leveraged
in subsequent code refactoring.
\end{acknowledgements}
